---
title: "gtsummary in quarto"
format:
  html:
    theme: pandoc 
    embed-resources: true
execute: 
  echo: false
---

**National Longitudinal Survey of Youth**

We’re using some data from the National Longitudinal Survey of Youth 1979, a cohort of American young adults aged 14-22 at enrollment in 1979. They continue to be followed to this day, and there is a wealth of publicly available data online. I’ve downloaded the answers to a survey question about whether respondents wear glasses, a scale about their eyesight with glasses, their (NLSY-assigned) race/ethnicity, their sex (presumably at birth, but unclear), their family’s income in 1979, and their age at the birth of their first child.

```{r}
#| warning: false
#| message: false

library(tidyverse)
library(gtsummary)
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd", 
               "id", "nsibs", "samp", "race_eth", "sex", "region", 
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"), 
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |> 
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes"))) 
```

# Customization of `tbl_summary()`

```{r}
#| label: tbl-one
#| tbl-cap: "This is a great table"

tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

@tbl-one shows the descriptive statistics

```{r}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```

```{r}
#| label: Last table

tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

```{r}
mean_wknd_sleep <- mean(nlsy$sleep_wknd, na.rm = T)

median_wknd_sleep <- median(nlsy$sleep_wknd, na.rm = T)
```

The mean number of hours of sleep on weekends is `r mean_wknd_sleep`. 

The mean number of hours of sleep on weekends is `r mean(nlsy$sleep_wknd, na.rm = T)`.

The median number of hours of sleep on weekends is `r median_wknd_sleep`. 

The median number of hours of sleep on weekends is `r median(nlsy$sleep_wknd, na.rm = T)`.


```{r}
#| label: tbl-descr
#| tbl-cap: "Descriptive statistics"
table1 <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir)) |> 
  add_overall(last = TRUE)
table1
```

```{r}
table1$table_body
```

```{r}
inline_text(table1, variable = "age_bir", column = "stat_1")
inline_text(table1, variable = "region_cat", level = "West",column = "stat_2")
	

```

```{r}
table2 <- tbl_summary(
  nlsy,
  include = c(sleep_wknd))
table2

table2$table_body

```

```{r}
sleep_wknd_median <- inline_text(table2, variable = "sleep_wknd", 
                         column = "stat_0")
```

The median weekend sleep hours is `r sleep_wknd_median`. 





